Manuscript

### 2019-01-24 manuscript_analysis.Rmd

Mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele_Length <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1_Length <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- ifelse(maf_df$Variant_Classification == "Silent", "Wildtype", "Mutant")
maf_df <- unique(maf_df)
saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

@@

```{r}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele_Length == 1 | is.na(ccle_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)
ccle_data_ptmuts$Dataset <- "CCLE"
ccle_data_ptmuts$Hugo_Symbol <- ordered(ccle_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele_Length == 1 | is.na(ctrp_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)
ctrp_data_ptmuts$Dataset <- "CTRP"
ctrp_data_ptmuts$Hugo_Symbol <- ordered(ctrp_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele_Length == 1 | is.na(gdsc_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)
gdsc_data_ptmuts$Dataset <- "GDSC"
gdsc_data_ptmuts$Hugo_Symbol <- ordered(gdsc_data_ptmuts$Hugo_Symbol, levels = g2p_genes)

@@

# Filter drug screen datasets for G2P drugs and genes
ccle_data_g2p_grid <- filter(ccle_data_ptmuts, Drug %in% g2p_drugs)
ccle_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_grid <- filter(ctrp_data_ptmuts, Drug %in% g2p_drugs)
ctrp_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_grid <- filter(gdsc_data_ptmuts, Drug %in% g2p_drugs)
gdsc_data_g2p_grid$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
```

```{r}
test_ccle <- ccle_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_ctrp <- ctrp_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_gdsc <- gdsc_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()

@@

```{r}
drug_summ <- data.frame(Drug = character(), Dataset_Count = integer(), stringsAsFactors = FALSE)

@@

# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "DepMap_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_crispr_18Q3 <- filter(ge_melt_18Q3, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
# ge_crispr_18Q4 <- filter(ge_melt_18Q4, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
crispr_data <- merge(crispr_data, ge_crispr_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# crispr_data <- merge(crispr_data, ge_crispr_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
saveRDS(crispr_data, "./../htscreens_giant_files/crispr_data_18Q4.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
crispr_data_ptmuts <- filter(crispr_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele_Length == 1 | is.na(crispr_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)

>>>>

Mutation calls (coding region, germline filtered):
```{r, eval=FALSE}
maf_raw <- read.delim("./data_munging/18Q4/depmap_18Q4_mutation_calls_20181114.csv.gz", header = TRUE, sep = ",", row.names = 1)
colnames(maf_raw)[colnames(maf_raw) == "Tumor_Sample_Barcode"] <- "CCLE_Name"
# Select columns
maf_df <- subset(maf_raw, select = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status_Nonsilent <- "Mutant"

saveRDS(maf_df, "./data_munging/rds/maf_df_18Q4.rds", compress = "xz")
```

@@

```{r, eval=FALSE}
ccle_data <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele == 1 | is.na(ccle_data_ptmuts$Reference_Allele), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)
ccle_data_ptmuts <- subset(ccle_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
ccle_data_ptmuts$Dataset <- "CCLE"
ccle_data_ptmuts$Hugo_Symbol <- ordered(ccle_data_ptmuts$Hugo_Symbol, levels = g2p_genes)
ccle_data_ptmuts <- unique(ccle_data_ptmuts)
saveRDS(ccle_data_ptmuts, "./data_munging/rds/ccle_data_18Q4_g2pgenes_ptmuts.rds", compress = "xz")

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele == 1 | is.na(ctrp_data_ptmuts$Reference_Allele), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)
ctrp_data_ptmuts <- subset(ctrp_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
ctrp_data_ptmuts$Dataset <- "CTRP"
ctrp_data_ptmuts$Hugo_Symbol <- ordered(ctrp_data_ptmuts$Hugo_Symbol, levels = g2p_genes)
ctrp_data_ptmuts <- unique(ctrp_data_ptmuts)
saveRDS(ctrp_data_ptmuts, "./data_munging/rds/ctrp_data_18Q4_g2pgenes_ptmuts.rds", compress = "xz")

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele == 1 | is.na(gdsc_data_ptmuts$Reference_Allele), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)
gdsc_data_ptmuts <- subset(gdsc_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
gdsc_data_ptmuts$Dataset <- "GDSC"
gdsc_data_ptmuts$Hugo_Symbol <- ordered(gdsc_data_ptmuts$Hugo_Symbol, levels = g2p_genes)
gdsc_data_ptmuts <- unique(gdsc_data_ptmuts)
saveRDS(gdsc_data_ptmuts, "./data_munging/rds/gdsc_data_18Q4_g2pgenes_ptmuts.rds", compress = "xz")
```

```{r}
ccle_data_ptmuts <- readRDS("./data_munging/rds/ccle_data_18Q4_g2pgenes_ptmuts.rds")
ctrp_data_ptmuts <- readRDS("./data_munging/rds/ctrp_data_18Q4_g2pgenes_ptmuts.rds")
gdsc_data_ptmuts <- readRDS("./data_munging/rds/gdsc_data_18Q4_g2pgenes_ptmuts.rds")

@@

# Filter drug screen datasets for G2P drugs and genes
ccle_data_g2p_grid <- filter(ccle_data_ptmuts, Drug %in% g2p_drugs)
ccle_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_grid <- filter(ctrp_data_ptmuts, Drug %in% g2p_drugs)
ctrp_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_grid <- filter(gdsc_data_ptmuts, Drug %in% g2p_drugs)
gdsc_data_g2p_grid$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Testing
test_ccle <- ccle_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_ctrp <- ctrp_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_gdsc <- gdsc_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()

@@

How many datasets screen each G2P drug?
```{r}
drug_summ <- data.frame(Drug = character(), Dataset_Count = integer(), stringsAsFactors = FALSE)

@@

# Cell line lineages
crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "DepMap_ID"), all.x = TRUE)
levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_crispr_18Q3 <- filter(ge_melt_18Q3, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
# ge_crispr_18Q4 <- filter(ge_melt_18Q4, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
crispr_data <- merge(crispr_data, ge_crispr_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# crispr_data <- merge(crispr_data, ge_crispr_18Q4, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
saveRDS(crispr_data, "./../htscreens_giant_files/crispr_data_18Q4.rds", compress = "xz")

# Takes a really long time to run (4+ hours)
# crispr_data_ptmuts <- filter(crispr_data, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
# crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele == 1 | is.na(crispr_data_ptmuts$Reference_Allele), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
# crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)
# crispr_data_ptmuts <- subset(crispr_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
# crispr_data_ptmuts <- unique(crispr_data_ptmuts)
# saveRDS(crispr_data_ptmuts, "./data_munging/rds/crispr_data_18Q4_ptmuts.rds", compress = "xz")
```

```{r}
crispr_data <- readRDS("./../htscreens_giant_files/crispr_data_18Q4.rds")
# crispr_data_ptmuts <- readRDS("./../htscreens_giant_files/crispr_data_18Q4_ptmuts.rds")
# 
# crispr_data_ptmuts <- filter(crispr_data, Hugo_Symbol %in% g2p_genes)
# crispr_data_ptmuts <- filter(crispr_data_ptmuts, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
# crispr_data_ptmuts$Keep_PtMut_Tests <- ifelse(crispr_data_ptmuts$Reference_Allele == 1 | is.na(crispr_data_ptmuts$Reference_Allele), TRUE, ifelse(crispr_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
# crispr_data_ptmuts <- filter(crispr_data_ptmuts, Keep_PtMut_Tests == TRUE)
# crispr_data_ptmuts <- subset(crispr_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
# crispr_data_ptmuts <- unique(crispr_data_ptmuts)
